---
title: "pun-v-assignment4"
author: "Vincent Pun"
date: "3/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This dataset summarizes 77 different brands of breakfast cereals, including calories, proteins, fats, and so on, and gives a 'rating' that indicates the overall nutritional value of the cereal.

```{r Load Data}
#LOAD DATA----

# Multidimensional Visualization Exercise: Breakfast Cereals
# 
# Description of the original data set from the American Statistical 
# Association (ASA): # http://stat-computing.org/dataexpo/1993.html
# 
# Install the library lgrdata for example datasets
# which includes most of the variables from the original ASA example 
# Documentation at
# http://search.r-project.org/R/library/lgrdata/html/cereals.html

library(lgrdata)

#https://www.rdocumentation.org/packages/lgrdata/versions/0.1.1/topics/cereals
data(cereals)

```

```{r Data Preparation}

#DATA PREPARATION----

#1----
# ignore the ratings column because it is undocumented
#double Health rating of the cereal (unknown calculation method).
cerealsall = cereals[,1:ncol(cereals)-1]

names(cerealsall)


# 2) use valid lowercase variable names---- 
names(cerealsall) = c("cereal","manufacturer","coldhot",
    "calories","protein","fat","sodium","fiber","carbo",
    "sugars","potass","vitamins")

#Summary
dim(cerealsall)
str(cerealsall)
print(summary(cerealsall)) 
```


```{r Subsets of Cereals}
#3) Subset of cereals----

# To simplify the problem, we could consider working with a 
# subset of cereals, such as only Kelloggs (manufacturer == "K") 
# and/or General Mills (manufacturer == "G")
# We do not do that here

# cereals with complete data
cerealscomplete = cerealsall[complete.cases(cerealsall),]
print(summary(cerealscomplete)) 

# only one hot cereal remaining, so focus on cold cereals 
#only C
#remove column 3 (coldhot)
cerealswork = 
    cerealscomplete[(cerealscomplete$coldhot == "C"),c(1,2,4:12)]
print(summary(cerealswork))  

```


```{r Rename observations}
str(cerealswork)
cerealswork
# Set max 8-character cereal names for annotation of visualizations
#Manually create abbreviation for each observation (name column)
cerealswork$name = c("100%Bran","100%NBrn","AllBran",                              
    "AllBranX","ACCherio","AplJacks",                           
    "Basic4",  "BranChex","BranFlak",                           
    "CapNCrun","Cheerios","CinTCrun",                 
    "Clusters","CocoPuff","CornChex",                             
    "CornFlak","CornPops","ContChoc",                         
    "COatBran","Crispix","CrispyWR",                
    "DbleChex","FrotLoop","FrosFlak",                        
    "FrosMWht","FFDWOats","FrutBran",                         
    "FrutPebb","GoldCrsp","GoldGram",                        
    "GNutsFlk","GrpeNuts","GGranPec",                    
    "HonyGOhs","HonyNChr","Honycomb",                            
    "JstRCrun","JstRFNut","Kix",                                   
    "Life","LkCharms","MusliRDA",
    "MusliRPP","MuslixCB","MGrnCher",                  
    "NutHonCr","NutGARai","NutgranW",                     
    "OatRaCrp","PstNRBrn","Prodct19",                            
    "PuffRice","PuffWhet","QuakOatS",                    
    "RaisBran","RaisNBrn","RaisSqur",                        
    "RiceChex","RiceKris","ShrdWeat",                        
    "ShrdWBrn","ShrdWSSz","Smacks",                               
    "SpecialK","StrwFWht","TotCornF",                     
    "TotRBran","TotWGran","Triples",                               
    "Trix","WhetChex","Wheaties","WhetHGld")

```


```{r}

# check the abbreviations (cereal vs name)
cat("\nName of Cereal (Abbreviation):\n")
for (i in 1:nrow(cerealswork)) 
    cat(cerealswork$cereal[i]," (",cerealswork$name[i],")\n", sep = "")

print(summary(cerealswork))  


```

***EDA***
```{r EDA}

#EDA----

library(DT)
library(dplyr)

unique(cerealswork$manufacturer)

#https://www.mrbreakfast.com/cereal_list_by_company.asp
#N = Nabisco
#Q = Quaker Oats
#K = Kelloggs
#G = General Mills
#R = Ralston
#P = Post
#A = A_HomestatFarm
cerealswork$manufacturername <- recode(cerealswork$manufacturer,"N" = "Nabisco","Q" = "Quaker_Oats", "K" = "Kelloggs", "G" = "General Mills", "R" = "Ralston","P" = "Post", "A" = "A_HomestatFarm")

datatable(data = cerealswork,
          rownames = FALSE,
          filter = "top",
          options = list(autoWidth = TRUE))

write.csv(cerealswork, file = "0_cerealswork.csv")
```

```{r Univariate EDA - geom_bar}

#identify dominant players in the market (Kelloggs and General Mills)
library(ggplot2)

g <- ggplot(cerealswork, aes(y = manufacturername))
g + geom_bar()
```

```{r Bivariate EDA - Calories}

#Histogram - Fiber

cerealswork %>% 
  ggplot(aes(x = fiber, fill = manufacturername)) +
    geom_histogram() +
   scale_x_continuous(name = "Fiber (grams)") +
    scale_y_continuous(name = "Count", limits = c(0, 11), breaks = seq(0, 11, 2)) +
    labs(fill = "Manufacturer", title = "Histogram - Fiber") +
    theme_minimal()

```
```{r Correlation Plot}

##GGally correlation plot to identify interesting variable relationships
library(GGally) #https://briatte.github.io/ggcorr/

ggcorr(cerealswork, geom = "blank", label = TRUE, hjust = 0.75) +
  geom_point(size = 10, aes(color = coefficient > 0, alpha = abs(coefficient) > 0.4)) +
  scale_alpha_manual(values = c("TRUE" = 0.25, "FALSE" = 0)) +
  guides(color = FALSE, alpha = FALSE)

```

```{r Corrplot - hierarchical clustering order}

#correlation plot with hierarchical clustering ordering
library(corrplot)
corrplot(cor(cerealswork[,3:11]),order='hclust', tl.cex = .9, tl.col = 'darkblue', addrect = 4, hclust.method = "complete")

```

```{r Bivariate EDA}
#Scatterplot - Carbs and Sugars
ggplot(cerealswork, aes(x=carbo, y=sugars, color=manufacturername)) + geom_point() + theme(legend.position=c(0,1), legend.justification=c(0,1)) + labs(color = "Manufacturer", title = "Scatterplot - Carbs and Sugar")

#Scatterplot - potass and fiber
ggplot(cerealswork, aes(x=potass, y=fiber, color=manufacturername)) + geom_point() + theme(legend.position=c(0,1), legend.justification=c(0,1)) + labs(color = "Manufacturer", title = "Scatterplot - Potassium and Fiber")

#Scatterplot - potass and fiber
ggplot(cerealswork, aes(x=potass, y=fiber, color=manufacturername)) + 
  labs(color = "Manufacturer", title = "Scatterplot - Potassium and Fiber")+
  geom_point(shape=1) +
    scale_colour_hue(l=50) + 
    geom_smooth(method=lm,se=FALSE)
```

```{r Skewness Kurtosis}

library(moments)


listnum <- c("calories","protein","fat","sodium","fiber","carbo","sugars","potass","vitamins")

for (i in listnum){
  print(paste(i,"  ",skewness(cerealswork[[i]])))
}

str(cerealswork)
```

Customers who are health conscious may be interested in comparing brands based on their sugar, calories, and carbs. 

```{r Normalize Data}
# At this point, there should be 73 cereals with complete data
# on 9 variables. The variables have meaningful magnitude but on
# various scales of measurement (some imprecise, some ordinal).
# Employ min-max normalization prior to analysis
minmaxnorm <- function(x) { return ((x - min(x)) / (max(x) - min(x))) }
minmaxdf <- as.data.frame(lapply(cerealswork[,3:11], FUN = minmaxnorm))

head(minmaxdf)
str(minmaxdf)

```
***Visualizations***
# Consider visualization methods such as metric multidimensional 
# scaling, nonmetric multidimensional scaling, biplots from 
# principal components, t-SNE, dendrograms from hierarchical 
# clustering, checkerboard plots from biclustering 
# (rows for cereals, columns for nutrition variables), 
# and self-organizing maps. Consider combinations of methods, 
# such as t-SNE plotting of groups of cereals identified by 
# K-means cluster analysis with a unique color for each cluster.

From G Search:
PCA assumes that the directions with the largest variances are the most “important” (i.e, the most principal). In the figure below, the PC1 axis is the first principal direction along which the samples show the largest variation. The PC2 axis is the second most important direction and it is orthogonal to the PC1 axis.

Loadings are interpreted as the coefficients of the linear combination of the initial variables from which the principal components are constructed. From a numerical point of view, the loadings are equal to the coordinates of the variables divided by the square root of the eigenvalue associated with the component.

```{r Screeplot}
#DATA VISUALIZATION----

library(psych)
library(lessR)
library(ggpubr)
library(factoextra)

#FACTOR ANALYSIS----
fa.parallel(minmaxdf, fa="both", n.iter=1000, show.legend=TRUE,main="Scree plot with parallel analysis")
```

```{r PCA}
#PCA----
#.rowNamesDF
.rowNamesDF(minmaxdf,make.names=TRUE) <-cerealswork$name

my.pca <- princomp(minmaxdf, scale. = TRUE)
summary(my.pca)

#view loadings
my.pca$loadings


```
```{r Screeplot 2}
fviz_eig(my.pca)
```

```{r PCA - Visualization}

fviz_pca_var(my.pca, 
             col.var = "contrib", 
             repel = TRUE, 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             title = "Principal Component Analysis",
             legend.title = "Contribution"
             )

biplot(my.pca, cex = .5,
       main="PCA Biplot")

```


```{r Factor Analysis}

#Factor Analysis----
#No Rotation, 3 factors
fa1<-fa(minmaxdf, nfactors=4, rotate="none", fm="pa")
summary(fa1)
fa1$loadings
```
```{r Heatmap and SemPaths}
library(gplots)
library(RColorBrewer)
library(semPlot)

heatmap.2(fa1$loadings,
          col=brewer.pal(9,"Blues"), trace="none", key=FALSE,
          Colv=FALSE, 
          cexCol = 1.2, cexRow = .7,
          dend="none", lwid=c(0.1,1), lhei=c(0.1,2)
          )

semPaths(fa1$loadings, what="est", residuals=FALSE,
         cut=0.5, posCol=c("white","darkgreen"), negCol=c("white","red"),
         edge.label.cex=1, nCharNodes=8, sizeMan = 7)
```


```{r Dimensionality Reduction}
minmaxdf_reduced <- minmaxdf[,!(names(minmaxdf) %in% c("sodium","vitamins"))]
minmaxdf_reduced
```

Using the reduced dataset produces an rsquare of 0.59, avg sillhouette 0.35
where using the regular normalized dataset produces rquare of 0.46, silhouette 0.43

```{r Kmeans}
#######################################################
### Create a Kmeans with 4 clusters with derived data #
#######################################################
set.seed(10)

clusterresults <- kmeans(x = minmaxdf_reduced,centers = 4)
names(clusterresults)

library(useful)
plot(clusterresults, data=minmaxdf_reduced)

dev.off()

#Between Sum of Squares / Total Sum of Squares Goodness to Fit 
#K-Means
rsquare <- clusterresults$betweenss/clusterresults$totss
rsquare

#K-Means, 3 clusters 

#MAC CAN'T HANDLET HIS
#plot(clusterresults, data = workdata_filtered_norm)

library(cluster)
dissE <- daisy(minmaxdf)
names(dissE)
dE2   <- dissE^2

sk2   <- silhouette(clusterresults$cluster, dE2)
str(sk2)
plot(sk2)

```

```{r  Plot kmeans}
#clusterresults cluster as dataframe
newdf <- as.data.frame(clusterresults$cluster)
pcadf <- as.data.frame(my.pca$scores)

write.csv(newdf, file = "clusterresults.csv")
write.csv(pcadf, file = "pca.csv")

dim(newdf)
dim(pcadf)
str(pcadf)

combdata <- cbind(newdf,pcadf)
head(combdata)

#xyplot of the 3 groups
library(lattice)
xyplot(Comp.2 ~ Comp.1, combdata, groups = clusterresults$cluster, pch= 15, main="K-means Clusters",xlab="Principal Component 1",ylab="Principal Component 2")

library(rgl)
bg3d("lightgrey")

#can see that within subgroups there is a split based on the third component, carbs 
plot3d(combdata$Comp.1,combdata$Comp.2,combdata$Comp.3,
       main = "K-Means Clusters Visualized on First 3 Principal Components",
       col=as.integer(cerealswork$manufacturer),
       size=4,
       radius=4)

?plot3d
text3d(combdata$Comp.1,combdata$Comp.2,combdata$Comp.3,rownames(minmaxdf2),color=clusterresults$cluster)

```

***Aggregate Data***
```{r Combined Data - Cluster Profiling}
################################################################
### Create a dataset with the original data with the cluster info----
### This will be useful for creating profiles for the clusters----
###############################################################

#cluster profiles + prepared data
combdata2 <- cbind(newdf, cerealswork)

#Rename column
require(reshape)
combdata2 <- rename(combdata2, c("clusterresults$cluster"="cluster"))
names(combdata2)

#Tabling Combdata2----

#Cluster 1 only has 3 observations
#Cluster 2 has 39 (largest group)
#Cluster 3 has 31

#table <- table(combdata2$cluster, combdata2$manufacturername)
#prop.table(table)

library(stats)
table_xtabs <- xtabs(~manufacturername + cluster, data=combdata2)
addmargins(table_xtabs)

prop_table_xtabs <- round((table_xtabs),2)
proptablextabsmargins <- as.data.frame.matrix(addmargins(prop_table_xtabs))
proptablextabsmargins$prop.g1 <- round(proptablextabsmargins$`1`/proptablextabsmargins$Sum,2)
proptablextabsmargins$prop.g2 <- round(proptablextabsmargins$`2`/proptablextabsmargins$Sum,2)
proptablextabsmargins$prop.g3 <- round(proptablextabsmargins$`3`/proptablextabsmargins$Sum,2)
proptablextabsmargins$prop.g4 <- round(proptablextabsmargins$`4`/proptablextabsmargins$Sum,2)

proptablextabsmargins

write.csv(proptablextabsmargins, file = "1_kmeans_proportion.csv")

#Aggregate columns by group----
#It is clear that cluster 1 is the niche healthy group
#Cluster 2 is less healthy, sweeter 
#Cluster 3 is healthy majority 

output_clusteraggregates <- aggregate(combdata2,by=list(byvar=combdata2$cluster), mean)
output_clusteraggregates

write.csv(output_clusteraggregates, file = "2_kmeans_mean.csv")

```


```{r t-SNE}
#default tsne
library(Rtsne)
set.seed(89)

tsne <- Rtsne(minmaxdf, dims = 2, perplexity = 10, verbose = TRUE, max_iiter=1000, learning=1000, check_duplicates=FALSE)

#visualizing
colors = rainbow(length(unique(clusterresults$cluster)))
names(colors) = unique(clusterresults$cluster)

par(mgp=c(2.5,1,0))

plot(tsne$Y, t='n', main = 'tSNE', xlab = 'tSNE dimension 1', ylab = 'tSNE dimension 2', cex.main = 2, cex.lab = 1.5)

text(tsne$Y, labels = combdata2$manufacturer, col = colors[clusterresults$cluster], cex=1, size=1)
?text
#t-SNE affirms that there are three distinct groups, found through kMeans
```

```{r 2-D MDS using minmaxdf2}
#MDS----

minmaxdf2 <- minmaxdf

#.rowNamesDF
.rowNamesDF(minmaxdf2,make.names=TRUE) <-cerealswork$name

d1 <- dist(minmaxdf2)

#cmdscale(), fit MDS 
?cmdscale
fit <- cmdscale(d1, eig=TRUE, k=4)

#plot solution
x <- fit$points[,1]
y <- fit$points[,2]

plot(x, y, xlab="Dimension 1", ylab="Dimension 2",
     main="Metric MDS", type="n")
text(x, y, labels = row.names(minmaxdf2), cex=.8)


```

```{r 3-D MDS using minmaxdf2}
minmaxdf2 <- minmaxdf

#.rowNamesDF
.rowNamesDF(minmaxdf2,make.names=TRUE) <-cerealswork$name

d1 <- dist(minmaxdf2)

#cmdscale(), fit MDS 
?cmdscale
fit <- cmdscale(d1, eig=TRUE, k=4)

#goodness of fit
fit$GOF

#plot solution
x <- fit$points[,1]
y <- fit$points[,2]
z <- fit$points[,3]

#3D
#https://stackoverflow.com/questions/52590410/display-point-labels-in-scatter3d

bg3d("lightgrey")

plot3d(fit$points[,1],fit$points[,2],fit$points[,3],
       main = "Multidimensional Scaling (3D)")
text3d(fit$points[,1],fit$points[,2],fit$points[,3],rownames(minmaxdf2),color=as.integer(cerealswork$manufacturer))

#play3d(spin3d(axis=c(0,1,1),rpm=3))

```

```{r MDS - look at eigenvalues, needs around 9 for a good fitting model}
#https://stats.stackexchange.com/questions/22019/how-to-calculate-the-r-squared-value-and-assess-the-model-fit-in-multidimensiona

r <- cmdscale(d1, eig=TRUE)

plot(cumsum(r$eig) / sum(r$eig),
     type="h", lwd=5, las=1, xlab="Number of dimensions", ylab=expression(R^2))

plot(r$eig, type="h", lwd=5, las=1, xlab="Number of dimensions", ylab="Eigenvalues")
```



```{r Checkerboard plot and biclustering}
#BICLUSTERING----
#blockcluster----
library(blockcluster)

minmaxdf3 <- as.matrix(minmaxdf_reduced)

set.seed(1)
?coclusterContinuous
blockobject <- coclusterContinuous(minmaxdf3, nbcocluster=c(2,3))

summary(blockobject)

plot(blockobject)
#biclust----
library(biclust)
x <- discretize(minmaxdf3, nof=3,quant=TRUE)
set.seed(2)
XMotifs <- biclust(x=as.matrix(x), method=BCXmotifs(),ns=5,nd=50,sd=3,alpha=0.01,number=100)

summary(XMotifs)
```

```{r Fuzzy Clustering}
#Fuzzy Clustering----
#https://www.datanovia.com/en/lessons/fuzzy-clustering-essentials/

#https://www.datanovia.com/en/lessons/fuzzy-clustering-essentials/cmeans-r-function-compute-fuzzy-clustering/

library(e1071)
set.seed(2)
cm <- cmeans(minmaxdf_reduced, centers=2, iter.max=100, dist="euclidean",m=2)

cm$membership
write.csv(cm$membership, file = "3_fuzzymembership.csv")

#Visualize Correlation
corrplot(cm$membership, is.corr=FALSE,tl.cex = .6, tl.col = 'darkblue')
?fviz_cluster

#plot clusters
fviz_cluster(list(data = minmaxdf, cluster=cm$cluster),
             ellipse.type="norm",
             ellipse.level=0.8,
             palette="jco",
             ggtheme=theme_minimal())
```


```{r Fuzzy Grouping}
fuzzy <- as.data.frame(cm$membership)

fuzzy$fuzzygroup <- ifelse(fuzzy$`1` > 0.7,"1",
                      ifelse(fuzzy$`2` > 0.7,2,3))
fuzzy

table(fuzzy$fuzzygroup)

round(prop.table(table(fuzzy$fuzzygroup)),2)

combdata3 <- cbind(fuzzy$fuzzygroup,cerealswork)

colnames(combdata3)[1] <- "group"

output_clusteraggregates_2 <- aggregate(combdata3[4:12],by=list(byvar=combdata3$group), mean)
output_clusteraggregates_2

write.csv(output_clusteraggregates_2, file = "5_fuzzy_mean.csv")

```

```{r Fuzzy Proportion Table}

#fuzzy table proportions 

table_xtabs_fuzzy <- xtabs(~manufacturername + group, data=combdata3)

#as.data.frame.matrix 
fuzzy <- as.data.frame.matrix(addmargins(table_xtabs_fuzzy))

fuzzy$prop.g1 <- round(fuzzy$`1`/fuzzy$Sum,2)
fuzzy$prop.g2 <- round(fuzzy$`2`/fuzzy$Sum,2)
fuzzy$prop.g3 <- round(fuzzy$`3`/fuzzy$Sum,2)
fuzzy

write.csv(fuzzy, file = "4_fuzzy_proportion.csv")

```



